---
title: "Homework 1"
author: "Artyom Muradyan"
date: "2025-02-05"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(readr)
```

1. Load the dataset. Check the first 5 rows.
```{r}
df <- read_csv("crime_data.csv")
head(df)

```
2. Identify columns with missing values and their respective counts. Drop columns where more than
50% of the data is missing (store this version as a new dataset).
```{r}
missing_counts <- colSums(is.na(df))
missing_counts
```
```{r}
columns_to_drop <- names(missing_counts[missing_counts > nrow(df) * 0.5])
df_new <- df %>% select(-all_of(columns_to_drop))
print(paste("Dropped columns:", paste(columns_to_drop, collapse=", ")))

```

3. Convert the DATE OCC column to a datetime format. Extract the year, month, and day into separate
columns. Create a new column for the hour using the TIME OCC column.

```{r}
df_new <- df_new %>%
  mutate(
    `DATE OCC` = as.Date(`DATE OCC`, format="%m/%d/%Y"),
    Year = format(`DATE OCC`, "%Y"),
    Month = format(`DATE OCC`, "%m"),
    Day = format(`DATE OCC`, "%d"),
    Hour = as.numeric(substr(sprintf("%04d", `TIME OCC`), 1, 2))
  )
```

4. Filter the dataset for crimes that occurred in 2023. Further filter crimes with the description
BURGLARY in the Crm Cd Desc column.

```{r}
df_new %>%filter(Year == 2023 & `Crm Cd Desc` == "BURGLARY")
```

5. Group the data by AREA NAME and calculate the total number of crimes and the average victim age.
Sort the results by total crimes in descending order.
```{r}

df_new %>%group_by(`AREA NAME`) %>%
  summarise(Total_Crimes = n(), Avg_Victim_Age = mean(`Vict Age`, na.rm = TRUE)) %>%
  arrange(desc(Total_Crimes))

```



 3.1. Group by Month and count crimes
```{r}
df_new %>% group_by(Month) %>% summarise(Total_Crimes = n())

```

3.2. Count crimes where a weapon was used
```{r}
# Weapon Used Cd column was dropped, hence using the original df not the cleaned one
df %>% filter(!is.na(`Weapon Used Cd`)) %>% summarise(Total_Crimes = n())

```

# 3. Group by Premis Desc and count crimes
```{r}
df_new %>% group_by(`Premis Desc`) %>% summarise(Total_Crimes = n())
```

4. Creating Severity Score
```{r}
# Weapon Used Cd column was dropped, hence using the original df not the cleaned one
df <- df %>%
  mutate(Severity_Score = ifelse(!is.na(`Weapon Used Cd`), 5, 0) +
                              ifelse(`Crm Cd Desc` == "BURGLARY", 3, 0) + 1)

df %>% group_by(`AREA NAME`) %>% summarise(Total_Severity_Score = sum(Severity_Score))

```


